---
title: "Homework 2"
author: Bin Yang
output: github_document
---

```{r setup}
library(tidyverse)
library(readxl)
```
## Problem 1

Read the Mr. Trashwheel dataset.

```{r}
trashwheel_df <- 
  read_xlsx("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
            sheet = "Mr. Trash Wheel", 
            range = cell_cols("A:N")) %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(
    sports_balls = round(sports_balls),
    sports_balls = as.integer(sports_balls)
  )
```

Read precipitation data for 2018 and 2017.

```{r}
precip_2018 <- 
  read_excel(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2018 Precipitation",
    skip = 1
  ) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2018) %>% 
  relocate(year)

precip_2017 <- 
  read_excel(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2017 Precipitation",
    skip = 1
  ) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2017) %>% 
  relocate(year)
```

Now combine annual precipitation. 

```{r}
month_df <-  
  tibble(
    month = 1:12,
    month_name = month.name
  )

precip_df <- 
  bind_rows(precip_2018, precip_2017)

left_join(precip_df, month_df, by = "month") 
```

This is dataset contains information from Mr. Trashwheel trash collector in Baltimore, MD. As trash enters the inner harbor, the trashweel collects that trash, and stores it in a dumpster. The dataset contains information on year, month, and trash collected, include some specific kinds of trash. There are a total of `r nrow(trashwheel_df)` rows in our final dataset. Additional data sheets include month precipitation data.  

## Problem 2

Read the NYC Transit Subway Entrance And Exit Data

```{r}
nyc_subway_df <- 
  read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv", skip = 1) %>% 
  janitor::clean_names() %>% 
  select(line: entry, vending, ada) %>% 
  mutate(entry = recode(entry, "YES" = TRUE, "NO" = FALSE ))
```
This data set contains information related to subway station in NYC, including line, station, name, station latitude / longitude, routes served, entry, vending, entrance type, exit, ADA compliance, staffing, staff hours, and entrance location. As a start, I cleaned the variables names, selected interested variables and changed the "entry" from character to logical type. After these steps, the resulting dataset has `r nrow(nyc_subway_df)` rows, and `r ncol(nyc_subway_df)` columns. The data however is not yet tidy. For next steps, I want to merge the route variables into one single column. 

The number of distinct stations.

```{r}
nyc_subway_df %>% 
  distinct(line, station_name) %>% 
  nrow()
```

The number of ADA compliant stations. 

```{r}
nyc_subway_df %>% 
  distinct(line, station_name, .keep_all = TRUE) %>% 
  filter(ada == TRUE) %>% 
  nrow()
```

The proportion of station entrances without vending allow entrance. 

```{r}
nyc_subway_df %>% 
  filter(entry == TRUE, vending == "NO") %>% 
  nrow() / nrow(filter(nyc_subway_df, vending == "NO"))
```

Reformat the dataset. 

```{r}
nyc_subway_df <- 
  nyc_subway_df %>% 
  mutate(across(contains("route"), as.character))
```

```{r}
nyc_subway_tidy_df <- 
  pivot_longer(
    nyc_subway_df,
    route1:route11,
    names_to = "route_number",
    names_prefix = "route",
    values_to = "route_name"
  )
```

The number of distinct stations that serves A train.

```{r}
nyc_subway_tidy_df %>% 
  filter(route_name == "A") %>% 
  distinct(line, station_name) %>% 
  nrow()
```

The number of ada compliant stations that serve A train.

```{r}
nyc_subway_tidy_df %>% 
  filter(route_name == "A", ada == TRUE) %>% 
  distinct(line, station_name) %>% 
  nrow()
```

## Problem 3 

Read in and tidy pol_months dataset. 

```{r}
pols_months_df <- 
  read_csv("./data/fivethirtyeight_datasets/pols-month.csv") %>% 
  separate(mon, sep = "-", into = c("year", "month", "day")) %>% 
  mutate(month = as.numeric(month)) %>% 
  mutate(month = month.abb[month]) %>% 
  pivot_longer(
    cols = starts_with("prez"),
    names_to = "president",
    values_to = "result"
  ) %>% 
  mutate(president = recode(president, "prez_dem" = "dem", "prez_gop" = "gop")) %>% 
  filter(result == 1) %>% 
  select(-day, -result)
```

Read in and tidy snp dataset.

```{r}
snp_df <- 
  read_csv("./data/fivethirtyeight_datasets/snp.csv") %>% 
  separate(date, sep = "/", into = c("month", "day", "year")) %>% 
  mutate(month = as.numeric(month)) %>% 
  mutate(month = month.abb[month]) %>% 
  select(-day) %>% 
  relocate(year)
```

Read in and tidy unemployment dataset. 

```{r}
uneply_df <- 
  read_csv("./data/fivethirtyeight_datasets/unemployment.csv") %>% 
    pivot_longer(
    cols = Jan:Dec,
    names_to = "month",
    values_to = "unemployment"
    ) %>% 
  rename(year = Year) %>% 
  mutate(year = as.character(year)) 
```

Merge snp and pols. 

```{r}
final_df  <- 
  left_join(pols_months_df, snp_df, by = c("year", "month"))

final_df <- 
  left_join(final_df, uneply_df, by =  c("year", "month"))
```

The “pols-month” dataset contains 822 observations of 9 variables regarding the information of politicians, including president, number of governors, representatives, and senators of each party from 1947 to 2015; The "snp" dataset contains 787 observations of 2 variables regarding the information of Standard & Poor’s stock market index (S&P) closing value from 1950 to 2015; The dataset “unemployment” contains 68 observations of 13 variables regarding the information of unemployment rate from 1948 to 2015.   

The merged dataset contains `r nrow(final_df)` observations and `r ncol(final_df)` variables. The final dataset contains information regarding number of politicians of each party, S&P closing values, and unemployment rate from 1947 to 2015.



